---
editor_options:
  markdown:
    wrap: 72
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/setup.R"))
```

## **Conclusion**

For the supervised learning model, we use sensitivity, specificity,
balanced accuracy, overall accuracy and kappa to compare the quality of
different models.

For the unsupervised learning model, we perform clustering and PCA
analysis to group a set of instances/features that share some common
characteristics. In addition, we also use the result from the PCA
analysis to perform a hybrid supervised/unsupervised learning model to
see if we can use PCA to improve the performance of the supervised
learning model.

### **Supervised learning**

> **One day of the week - Unbalanced data**

According to table1, table2, the overall accuracy and kappa of decision
tree and multinomial regression are very low. What's more, there is a
serious unbalance between sensitivity and specificity. The quality of
both models is pretty similar and needs to be improved.

Accuracy: decision tree:**O.211** multinomial logistic
regression:**O.213**

kappa:decision tree:**O.016** multinomial logistic regression:**O.01**

***Table 1 - Decision Tree***

```{r, echo = FALSE, message = FALSE, warning=FALSE}
by_day_unbal_tree1 <- as.data.frame(c1$byClass) %>% select(1,2,11)
by_day_unbal_tree2 <- data.frame(cbind(Accuracy=c1$overall[1],
                                     Kappa=c1$overall[2]))
row.names(by_day_unbal_tree2) <- "overall"

knitr::kable(
  list(by_day_unbal_tree2, by_day_unbal_tree1),
  caption = 'decision tree',
  booktabs = TRUE, valign = 't')
```

***Table 2 - Multinomial Logistic Regression***

```{r}
by_day_unbal_mlr1 <- as.data.frame(c5$byClass) %>% select(1,2,11)
by_day_unbal_mlr2 <- data.frame(cbind(Accuracy=c5$overall[1],
                                     Kappa=c5$overall[2]))
row.names(by_day_unbal_mlr2) <- "overall"
 
knitr::kable(
  list(by_day_unbal_mlr2, by_day_unbal_mlr1),
  caption = 'multinomial logistic regression',
  booktabs = TRUE, valign = 't')
```

> **One day of the week - balanced data**

According to table3 and table4, after we balanced the data and do the
cross-validation, the difference between sensitivity and specificity is
smaller. Even though the accuracy is a little lower, the kappa of both
decision tree and multinomial logistic regression are higher.

Accuracy: decision tree:**O.195** multinomial logistic
regression:**O.176**

kappa:decision tree:**O.035** multinomial logistic regression:**O.03**

***Table 3 - with cross-validation: Decision Tree***

```{r}
by_day_bal_tree1 <- as.data.frame(c2$byClass) %>% select(1,2,11)
by_day_bal_tree2 <- data.frame(cbind(Accuracy=c2$overall[1],
                                     Kappa=c2$overall[2]))
row.names(by_day_bal_tree2) <- "overall"

knitr::kable(
  list(by_day_bal_tree2, by_day_bal_tree1),
  caption = 'with cross-validation: decision tree',
  booktabs = TRUE, valign = 't')
```

***Table 4 - Multinomial Logistic Regression***

```{r}
by_day_bal_mlr1 <- as.data.frame(c6$byClass) %>% select(1,2,11)
by_day_bal_mlr2 <- data.frame(cbind(Accuracy=c6$overall[1],
                                     Kappa=c6$overall[2]))
row.names(by_day_bal_mlr2) <- "overall"
 
knitr::kable(
  list(by_day_bal_mlr2, by_day_bal_mlr1),
  caption = 'multinomial logistic regression',
  booktabs = TRUE, valign = 't')
```

> **Weekdays and Weekend-Balanced data with Cross-Validation**

According to the table5, the balanced accuracy and kappa is higher after
we aggregate the days in a week into weekday and weekends. We will
finally choose logistic regression model for it has highest balanced
accuracy and kappa.

***Table 5 - The scores of three models***

```{r}
final_table <- data.frame(cbind(Sensitivity=c(c3$byClass[1],
                                              c4$byClass[1],
                                              c7$byClass[1]),
                                Specifity=c(c3$byClass[2],
                                              c4$byClass[2],
                                              c7$byClass[2]),
                                Accuracy=c(c3$overall[1],
                                           c4$overall[1],
                                           c7$overall[1]),
                                Balanced_accuracy=c(c3$byClass[11],
                                              c4$byClass[11],
                                              c7$byClass[11]),
                                Kappa=c(c3$overall[2],
                                           c4$overall[2],
                                           c7$overall[2])))
row.names(final_table) <- c("Decison Tree","Random Forest","Logistic Regression")

kable(final_table, caption = "The scores of three models (Decison Tree, Random Forest, Logistic Regression)") %>%
  kable_styling(bootstrap_options = "bordered") 
```

### **Unsupervised learning**

> **Clustering**

Clustering is a method that divides categorized or uncategorized data
into similar groups or clusters. We used 2 methods for clustering,
including Hierarchical clustering and Partitioning, and chose the number
of clusters from statistics. We found that the number of clusters highly
depend on the statistics chosen. For example, for the Partitioning
method, the optimal number of clusters from Gap statistic is 13, while
the optimal number from Silhouette is only 2.

For Hierarchical clustering, we chose 3 as an optimal number of
clusters. The main characteristics of each cluster are, for example,
high "produce" for cluster 1, high "snacks" and "beverages" for cluster
3. The average numbers of purchases from cluster 1 and 3 are
significantly higher than cluster 2 in every department.

For Partitioning methods, we chose 2 as an optimal number of clusters
for both K-mean and PAM models. For K-mean, we found that cluster 2 has
higher average numbers of purchases than cluster 1 in every department.
This means that cluster 2 in K-mean has similar characteristics as
cluster 1 and 3 from the Hierarchical clustering. For PAM, the results
are very close to the results from K-mean.

> **PCA**

To derive the benefits of PCA and to test our hypothesis that grouping
aisles reflects purchasing customer behavior better than grouping
departments, we perform PCA by aisles. Non-scaled PCA is our focus to
set the environment in line with the supervised learning analysis. As
for the result, component1 to component28 (out of 134 components) can
explain 75.5% of the total variation of the data. Fresh vegetable, fresh
fruits and packaged vegetables fruits are the top three most variance
and contributors to component1 and 2. Scaled PCA is also performed and
we need 93 components out of 134 components to explain at least 75% of
the total variation. Thus, we cannot derive the benefits of PCA and it
also shows that each aisle is not correlated. Afterward, hybrid
supervised and unsupervised learning approach grabs our attention. The
1-28 components are applied to the best model from supervised learning
approach which is the logistic regression model. However, we find that the hybrid method doesn't improve the accuracy of the prediction.

## **Limitation**

The data set provided by Instacart contains only a few variables,
including products purchased per order, product description, day of the
week of orders, and hours of orders. We found that this information is
still not sufficient to create predictive models with high accuracy for
commercial uses. However, Amazon is an ideal case study showing that,
with high quality and comprehensive data set, it's even possible to use
machine learning for a predictive analysis to anticipate and ship
products before customers actually order it.

We also find the limitation when running machine learning functions in R with too many observations. In our study, we focus only on the training data set, which contains over 100k observations and do not include the prior data set, which represents all historical orders before the training set and contains over a million observations. This is because the execution time is extremely long and there are errors regularly when we run the models with all historical data (prior + training).

## **Further study**

Machine learning and Artificial intelligence have play a key role in the
boom of the e-commerce industry. There are a many applications that we
can apply machine learning techniques in this industry. For example,
Purchase and Repurchase prediction for predicting what orders customers
will purchase next, Recommendation system for recommending products that
users might would like to purchase. This system is very important behind
the success of Amazon and even Netflix. There are also other
applications such as Fraud prediction and Marketing campaign. However,
in order to build these models with sufficient accuracy for commercial
use, it's important to we need more data such as customer personal
information, browsing history, click history to get a better
understanding of customer behaviors.

## **References**

[Ivo Bernardo, Classification Decision Trees, Easily Explained, Aug 30,
2021](https://towardsdatascience.com/classification-decision-trees-easily-explained-f1064dde175e)

[Saikumar Talari, Random Forest® vs Decision Tree: Key Differences,
February 18,
2022](https://www.kdnuggets.com/2022/02/random-forest-decision-tree-key-differences.html#:~:text=The%20critical%20difference%20between%20the,work%20according%20to%20the%20output.)

[Wikipedia, Multinomial logistic regression, 23 August
2021](https://en.wikipedia.org/wiki/Multinomial_logistic_regression)

S. Walusala W, R. Rimiru, C. Otieno, A hybrid machine learning approach
for credit scoring using PCA and logistic regression, International
journal of computer, ISSN 2307-4523.

[Jeremy Stanley, 3 Million Instacart Orders, Open Sourced, May 3,
2017](https://tech.instacart.com/3-million-instacart-orders-open-sourced-d40d29ead6f2)

[Bigcommerce Blog, Nick Shaw, Ecommerce Machine Learning: AI's Role in
the Future of Online
Shopping](https://www.bigcommerce.com/blog/ecommerce-machine-learning/#machine-learning-has-developed-over-the-years)

[The Economic Time, Amazon may predict and ship your order before you
place it, Jan 27,
2014](https://economictimes.indiatimes.com/tech/internet/amazon-may-predict-and-ship-your-order-before-you-place-it/articleshow/29445699.cms)
